\section{LLGR Stale Route Timer Experiment}

\subsection{Purpose}

This experiment was performed to observe stale-route handling and withdrawal timer behaviour with Long-Lived Graceful Restart. The goal was to check whether a route learned from a failed BGP peer remains available temporarily and is removed after the configured timer period.

\subsection{Experiment Setup}

The route tested was:

\begin{verbatim}
10.0.93.0/24
\end{verbatim}

This prefix is originated from the ISP/upstream side behind \texttt{hpe-r9}. The observing router was \texttt{hpe-r1}.

For the experiment, the timers were temporarily reduced:

\begin{table}[H]
\centering
\small
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{|p{6cm}|p{5cm}|}
\hline
\textbf{Timer} & \textbf{Temporary Value} \\
\hline
Graceful Restart time & 5 seconds \\
\hline
LLGR stale time & 10 seconds \\
\hline
Expected total retention window & About 15 seconds \\
\hline
\end{tabular}
\caption{Temporary GR and LLGR timer values}
\end{table}

After applying the temporary timer configuration, the BIRD daemon on \texttt{hpe-r9} was stopped. The route table on \texttt{hpe-r1} was then sampled repeatedly to check when the route disappeared.

\subsection{Measurement Result}

\begin{table}[H]
\centering
\small
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{|p{6cm}|p{5cm}|}
\hline
\textbf{Measurement} & \textbf{Result} \\
\hline
Target prefix & \texttt{10.0.93.0/24} \\
\hline
GR timer & 5 seconds \\
\hline
LLGR stale timer & 10 seconds \\
\hline
Last time route from \texttt{hpe-r9} was present & 14259 ms \\
\hline
First time route from \texttt{hpe-r9} became missing & 15350 ms \\
\hline
LLGR stale marker observed & Yes, shown as \texttt{(100s)} and \texttt{BGP.community: (65535,6)} \\
\hline
\end{tabular}
\caption{LLGR stale route timer result}
\end{table}

\subsection{Observation}

The route did not disappear immediately after the peer-side failure. Instead, it remained present for approximately 14.259 seconds and disappeared around 15.350 seconds.

This closely matches the configured timing expectation of 5 seconds of Graceful Restart time plus 10 seconds of LLGR stale time.

The literal text \texttt{LLGR\_STALE} was not printed by BIRD. However, the stale route marker was visible as \texttt{(100s)} and the BGP community \texttt{(65535,6)}. The community \texttt{65535:6} corresponds to the LLGR stale community. Therefore, this experiment confirms both stale-route retention timing and visible LLGR stale marking.

\subsection{Conclusion}

This experiment demonstrates stale-route retention and withdrawal timer behaviour. With shortened GR and LLGR timers, the route remained temporarily available, was marked stale using \texttt{(100s)} and \texttt{BGP.community: (65535,6)}, and was removed after approximately the expected timer duration. This satisfies the stale-route handling and withdrawal-timer observation requirement.
